---
title: "Flight Analysis"
output: html_notebook
---

An examination of flight patterns from data included in the nycflights13 package. 

```{r}
install.packages('nycflights13')
```

```{r}
library(nycflights13)
library(dplyr)
library(tidyverse)
library(ggplot2)
```

Exploratory analysis of flights dataframe.

```{r}
str(flights)
dim(flights)
sum(is.na(flights))
```

We can see that their are 19 variables and 336,776 rows in the dataframe. Their are numerous entries, 46595, containing NA values which will be addressed throughout analysis. 

```{r}
meanDis <- mean(flights$distance, na.rm=TRUE)
cat("The average distance travelled during all flights that supplied distance data is", meanDis)
```

We can filter flights similarly to the way would using SQL to comprise a list of flights operated by a particular carrier. Flights operated by Delta, American Airlines, and United Airlines.
```{r}
AUDFlights <- filter(flights, carrier == 'AA' | carrier == 'UA' | carrier == 'DL')
```

This can be accomplished with piping using the alternative line of code below.

```{r}
AUDFlightsTidy <- filter(flights, carrier %in% c('AA', 'UA', 'DL'))
```

Side by side comparison of the two datasets:

```{r}
head(AUDFlights)
head(AUDFlightsTidy)
tail(AUDFlights)
tail(AUDFlightsTidy)
AUDFlights[323,]
AUDFlightsTidy[323,]
```

As you can see the first 10 rows and columns of each dataset are identical. Selecting a random row, 323, in the dataframe revealed identical data entries. 

We can answer questions about the performance of each airline using the newly created dataset. 

Q1. Which carrier had the longest departure delay on Christmas day?

```{r}
chrisDay <- filter(AUDFlights, month == 12 & day == 25)
chrisDayT <- filter(AUDFlightsTidy, month == 12 & day == 25)
CDDelay <- arrange(chrisDay, desc(dep_delay))
CDDelay[1,10]
```
Delta had the longest departure delay on Christmas day! We can pull the actual value representing the length of the departure delay in the code below.
```{r}
CDDelay[1,6]
```

A client may be interested in how many flights were operated on Christmas day. We can simply count the total entries in the dataframe. 
```{r}
count(AUDFlights)
```

It may be important to investigate how often airlines are leaving late.

```{r}
deltaLate <- filter(AUDFlights, carrier == 'DL' & dep_delay > 0)
count(deltaLate)
aALate <- filter(AUDFlights, carrier == 'AA' & dep_delay > 0)
uALate <- filter(AUDFlights, carrier == 'UA' & dep_delay >0)
tabLate <- table(deltaLate)
tabLate
```


When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

