---
title: "Stroke Predictive Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Dataset source: https://www.kaggle.com/fedesoriano/stroke-prediction-dataset 
Attribute Information
1) id: unique identifier
2) gender: "Male", "Female" or "Other"
3) age: age of the patient
4) hypertension: 0 if the patient doesn't have hypertension, 1 if the patient has hypertension
5) heart_disease: 0 if the patient doesn't have any heart diseases, 1 if the patient has a heart disease
6) ever_married: "No" or "Yes"
7) work_type: "children", "Govt_jov", "Never_worked", "Private" or "Self-employed"
8) Residence_type: "Rural" or "Urban"
9) avg_glucose_level: average glucose level in blood
10) bmi: body mass index
11) smoking_status: "formerly smoked", "never smoked", "smokes" or "Unknown"*
12) stroke: 1 if the patient had a stroke or 0 if not
*Note: "Unknown" in smoking_status means that the information is unavailable for this patient
```{r}
#import csv
strokeData <- read.csv('/Volumes/Starbuck/BioinforPy/HealthCareDataR/healthcare-dataset-stroke-data.csv', stringsAsFactors = TRUE)
```


```{r}
#get a list of the column names
names(strokeData)
```
```{r}
#Determine which predictors are statistically significant in determining whether a stroke will occur
#create data frame 
strokeDf <- data.frame(strokeData)
#view first 10 rows
head(strokeDf, 10)
```


```{r}
#get a list of column names
names(strokeDf)
```
Let's start by examining the impact of hypertension on stroke incidence for the patients involved. 
```{r}
#Recall that a stroke is denoted by the number '1' 
cat("Variables Prior to Transformation:\n")
str(strokeDf)
#convert the stroke variable into a 2 level factor
strokeDf$stroke <- as.factor(strokeDf$stroke)
strokeDf$hypertension <- as.factor(strokeDf$hypertension)
cat("Variables After Two-Level Factor Conversion of Stroke:\n")
str(strokeDf)
```

```{r}
strokeMod <- glm(stroke~hypertension, data = strokeDf, family='binomial')
summary(strokeMod)
```
This logistic Regression model provides the following insights:
```{r}
b0 <- coef(strokeMod)
cat("The intercept coefficient value is", round(b0[1],3), "which represents the odds of people without hypertension having a stroke.")
```
Let's make predictions with based on the data we have.
```{r}
strokenohtn <- data.frame(hypertension="0")
nohtnpred <- predict(strokeMod, strokenohtn, type='response')
pctnohtn <- nohtnpred*100
cat("Patients without hypertension, within this group, have a", round(nohtnpred,5), "probability of stroke which indicates a", round(pctnohtn, 3), "% chance of having a stroke.\nRelatively good odds of avoiding a life altering event such as a stroke.")
```
```{r}
strokeWhtn <- data.frame(hypertension='1')
htnPred <- predict(strokeMod, strokeWhtn, type='response')
pctWhtn <- htnPred*100
cat("Hypertensive patients, within this group, have a", round(htnPred,5), "probability of stroke which indicates a", round(pctWhtn, 3), "% chance of having a stroke.\nThis signals an increase in risk for patients diagnosed with hypertension.")
```
Let's take a look at the impact of heart disease on the risks of stroke in the patients included in this study.

```{r}
hDmodStrokeMod <- glm(stroke~heart_disease,data = strokeDf, family='binomial')
summary(hDmodStrokeMod)
```
```{r}
noHdpred <- data.frame(heart_disease="0")
noHD <- predict(hDmodStrokeMod, noHdpred, type='response')
pctnoHd <- noHD*100
cat("Patients without heart disease, within this group, have a", round(noHD,5), "probability of stroke which indicates a", round(pctnoHd, 3), "% chance of having a stroke.")
```

```{r}
hasHDdat <- data.frame(heart_disease="1")
yesHDpred <- predict(hDmodStrokeMod, hasHDdat, type='response')
pcthasHd <- yesHDpred*100
cat("Patients with heart disease, within this group, have a", round(yesHDpred,5), "probability of stroke which indicates a", round(pcthasHd, 3), "% chance of having a stroke.\nThis signals an increase in risk for patients diagnosed with hypertension.")
```

